---
title: "Streaming Service Movie Rater"
author: "Chris Shin, Dylan Clark-Boucher, Grace Bosma"
date: "Due: 12/22/2020"
output: pdf_document
---

```{r, include= FALSE}
library(pacman)
pacman::p_load(tidyr, dplyr, ggplot2, gridExtra)
gbtheme <-theme_light() + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.ticks = element_line(color = "grey60"), plot.title = element_text(hjust = 0.5), text = element_text(size = 10, face = "bold"))
load(file = "movies.rda")
```

# Objective 
The purpose of this project was to both model and visualize IMDb ratings of movies and films currently available for streaming on various online platforms

# Introduction 
According to Statista, approximently 74% of homes have access to at least one streaming service in 2019 compared to about 52% in 2015. Additionally, as a result of the COVID-19 pandemic, many consumers have either purchased an additional streaming service subscription or increased time spent streaming. It is clear that movie and TV streaming services are increasing in popularity and are integrated into many of our day to day routines. Just last year, the movie and television streaming industry was valued $42.60 billion USD. Information on the popularity of movies available for streaming is critical to streaming platforms like Netflix, Disney+, Hulu to maintain customer interest in their platform rather than their competitors. 

# Methods
The data for this analysis was obtained from either Kaggle or International Movie Database (IMDb) websites. Data regarding streaming service platform and ratings from IMDb and Kaggle was obtained from Kaggle whereas information on actors involved in movies were obtained from IMDb. 

The Kaggle Movie data set (link:  https://www.kaggle.com/ruchi798/movies-on-netflix-prime-video-hulu-and-disney) was downloaded on 12/4/2020. This dataset contained information on movies available on various streaming services as of May 22, 2020 from US locations. In the dataset there were 16,631 movie titles (entries) and 17 variables including Title, Year, Age, IMDb (rating from IMDb, which is aggregated from IMDb registered users who cast a vote from 1 to 10 per title), Rotten.Tomatoes (TOMATOMETER score from Rotten Tomatoes), Netflix (whether it was available on Netflix), Hulu (whether it was available on Hulu), Prime.Video (whether it was available on Amazon Prime Video), Disney (whether it was available on Disney+), Type (whether it was a movie), Directors, Genres, Country, Languages, and Runtime. During cleaning, movies that did not match to titles on IMDb were excluded. The final movie data set consisted of 11448 titles.  

Data made publicly available on the website IMDb is updated daily and available for download from the website https://www.imdb.com/interfaces/ . The dataset "name.basics.tsv.gz" was downloaded from IMDb pm 12/4/2020 and used in this project. This data set contained information on actors and held information on 6,414,148 movie production team members, four "Known For" film names associated with each actor or actress, and additional roles the actor or actress held (i.e. writer, producer, ...). During cleaning, those that did not hold an actor/actress role or had no "known for" titles were excluded from analyses. The final, cleaned IMBd actors dataset held information on 2,426,996 actors or actresses. 

Data for prediction using linear regression and random forest required complete cases, and the variables we chose to include were age_rating (7+, 13+, 16+, 18+, or all), hulu (whether it was available on Hulu), disneyplus (whether it was available on Disney+), netflix (whether it was available on Netflix), amazonprime (whether it was available on Amazon Prime), year (the year the movie was released), rt_score (TOMATOMETER score from Rotten Tomatoes), imdb_score (rating from IMDb, which is aggregated from IMDb registered users who cast a vote from 1 to 10 per title), actor_score1, actor_score2, and actor_score3. 

- include something about any new variables created?

# Prediction
Based on ratings data mentioned above, a flexible model was constructed for predicting overall ratings on a scale of 0 to 5 stars. These predictions depended on a user-defined set of covariates, including such factors as a production's streaming service(s), target age group, release year, Rotten Tomato's Average Tomatometer score, and others.

- lm model

# Vizualization
```{r, echo = FALSE, warning=FALSE}
imdb_p <- movies_clean %>% select(year, imdb_score) %>% ggplot(aes(year, imdb_score)) + geom_point(col = 'plum4')+ gbtheme + ggtitle("IMDB Rating Over Time") + ylab("IMDB Score")

rt_p <- movies_clean %>% select(year, rt_score ) %>% ggplot(aes(year, rt_score)) + geom_point(col = 'steelblue4')+ gbtheme + ggtitle("Rotten Tomato Rating Over Time") + ylab("Rotten Tomato Score")

gridExtra::grid.arrange(imdb_p, rt_p, ncol=2)
```

- not married to these plots just figured we should show something 

# RShiny Application
An [RShiny Application](https://dylanclark-boucher.shinyapps.io/movie_rater_shiny/) was created using our data set to display our results and can be accessed at:  
```https://dylanclark-boucher.shinyapps.io/movie_rater_shiny/```

There are four tabs in the upper left that allows you to navigate through the application; Ratings Filter Tool, Ratings by Category, Score Prediction, and About. The Ratings filter tool presents a histogram and box plot of movie ratings and allows the user to filter based on recommended age of target audience, streaming service, Rotten Tomato Score, and release year and select a preferred choice of graph color and bin width. The Ratings By Category page presents three tabs on the left side bar. The first, Streaming Service, presents the distributions of movie ratings for each streaming service in a box plot format. The second, Age Rating, presents the distributions of movie ratings for by age group in a box plot format. The third, presents a scatter plot of IMBd movie ratings plotted against Tomatometer Score. Below the scatter plot the user is again able to filter but age group and streaming service. THe third tab, Score Prediction, [FILL IN LATER XXX]. The final tab, About, briefly summarized the background, app, model and contributors.  

- potentially integrate an example here?  

# Discussion 
- conclusion from lm model 
  - what variables are important
  - did we predict well 
- future direction

# Contributions 
Our group collaborated well and distributed work evenly among all members. While all members participated in searching for potential data sets, organizing, drafting and proofreading, Dylan created the RShiny app and organized a shared folder on the biostat computing cluster, Chris cleaned the Kaggle and movie rating data sets and Grace cleaned the actors data set and created a "cast score" to be used in model. 

- be sure to update and include model
